---
title: "Friday the 13th EDA"
author: "Luis de Sousa"
date: "October 4, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.path = "imgs/"
)
```

# U.S. Births

This folder contains data behind the story [Some People Are Too Superstitious To Have A Baby On Friday The 13th](http://fivethirtyeight.com/features/some-people-are-too-superstitious-to-have-a-baby-on-friday-the-13th/).

`us_births_2000-2014.csv` contains U.S. births data for the years 2000 to 2014, as provided by the Social Security Administration.

The data have the following structure:

Header | Definition
---|---------
`year` | Year
`month` | Month
`date_of_month` | Day number of the month
`day_of_week` | Day of week, where 1 is Monday and 7 is Sunday
`births` | Number of births

There are 3 main goals for DataExplorer:

# 1. Exploratory Data Analysis (EDA)

```{r load}
#load the library
library(DataExplorer)
#load input dataset
us_births <- read.csv('us_births_2000-2014.csv', header = T)
us_births$year<-as.factor(us_births$year)
us_births$month<-as.factor(us_births$month)
us_births$date_of_month<-as.factor(us_births$date_of_month)
us_births$day_of_week<-as.factor(us_births$day_of_week)
```

## Variables
```{r variables}
plot_str(us_births)
```

## Snapshot
```{r snapshot}
introduce(us_births)
```

## Missing Values
```{r missing values}
plot_missing(us_births)
```

## Histogram
```{r histogram}
plot_histogram(us_births)
```

## Density plot
```{r density plot}
plot_density(us_births)
```

## Correlation Plot
```{r correlation plot}
plot_correlation(us_births, type = 'continuous')
```

## Barplots
```{r barplots}
#No Discrete Features
#plot_bar(us_births)
```

## Boxplot
```{r boxplot}
plot_boxplot(us_births, by = "date_of_month") 
```

## Scatterplot
```{r scatterplot}
plot_scatterplot(us_births, by = "date_of_month", size = 0.5)
```

## Principle Component Analysis
```{r}
pca_df <- na.omit(us_births[, c("year","month","date_of_month")])
plot_prcomp(pca_df)
```

# 2. Feature Engineering
## Replace missing values
```{r}
final_df <- set_missing(us_births, list(0L, "unknown"))
plot_missing(final_df)
```

## Group sparse categories
```{r}
final_df <- group_category(data = us_births, feature = "month", threshold = 0.2, update = TRUE)
plot_bar(final_df$month)
```

## Dummify data (one hot encoding)
```{r}
plot_str(
    list(
        "original" = us_births,
        "dummified" = dummify(us_births)
    )
)
```

## Drop features
```{r}
#identical(
#  drop_columns(final_data, c("dst_origin", "dst_dest", "tzone_dest")),
#  drop_columns(final_data, c(34, 41, 42))
#)
```

# 3. Data Reporting
## Create Report
```{r create report, echo=TRUE, results='hide', message=FALSE}
create_report(us_births)
```